---
title: "Mesh sensitivy analysis per species and region"
date: "`r Sys.Date()`"
output: 
  rmdformats::downcute:
  highlight: zenburn
html_document :
  toc: true
toc_float: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = getwd())                                # garde le working dir actif
knitr::opts_knit$set(envir = globalenv())                               # utilise l'env global
knitr::opts_chunk$set(echo = FALSE, message=FALSE,warning=FALSE)
```


<br/>
 
```{r echo=FALSE}

print(sp_scientific)

```
<br/>
<br/>

# 1. Sanity results for meshes configurations
<br/>
```{r echo=FALSE, message=FALSE, warning=FALSE}

conv_simple <- if (!is.null(sanity_simple)) names(sanity_simple$cutoffs_converged_models) else character(0)
fail_simple <- if (!is.null(sanity_simple)) names(sanity_simple$cutoffs_sanity_failed) else character(0)

conv_barrier <- if (!is.null(sanity_barrier)) names(sanity_barrier$cutoffs_converged_models) else character(0)
fail_barrier <- if (!is.null(sanity_barrier)) names(sanity_barrier$cutoffs_sanity_failed) else character(0)

conv_boundary <- if (!is.null(sanity_boundary)) names(sanity_boundary$cutoffs_converged_models) else character(0)
fail_boundary <- if (!is.null(sanity_boundary)) names(sanity_boundary$cutoffs_sanity_failed) else character(0)


sanity_results_df <- dplyr::bind_rows(

  tibble(cutoff = region_params$cutoff_values, 
         config = "simple", 
         status = dplyr::case_when(
           paste0("cutoff_", cutoff) %in% conv_simple ~ "converged",
           paste0("cutoff_", cutoff) %in% fail_simple ~ "failed")),
  tibble(cutoff = region_params$cutoff_values, 
         config = "barrier",
         status = dplyr::case_when(
           paste0("cutoff_", cutoff) %in% conv_barrier ~ "converged",
           paste0("cutoff_", cutoff) %in% fail_barrier ~ "failed")),
  tibble(cutoff = region_params$cutoff_values,
         config = "boundary",
         status = dplyr::case_when(
           paste0("cutoff_", cutoff) %in% conv_boundary ~ "converged",
           paste0("cutoff_", cutoff) %in% fail_boundary ~ "failed")))

sanity_results_df$config <- factor(sanity_results_df$config,
                                   levels = c("boundary", "barrier", "simple"))

ggplot(sanity_results_df, aes(x = factor(cutoff), y = config, fill = status)) +
  geom_tile(color = "white") +
  scale_fill_manual(values = c(converged = "#2ca25f", failed = "#ffb703")) +
  labs(x = "Cutoff", y = "Meshes configuration", fill = "Model status") +
  theme_minimal()

```

<br/>  
<br/>

# 2. Estimates and metrics for meshes configurations
<br/>

```{r echo=FALSE, fig.height=6, fig.width=10, message=FALSE, warning=FALSE}

all_params %>%
  mutate(config = case_when(!barrier & !boundary ~ "simple",
                            !barrier &  boundary ~ "boundary",
                            barrier & !boundary ~ "barrier")) %>%
  filter(term %in% c("(Intercept)", "range", "sigma_O")) %>%
  rename(metric = term, value = estimate) %>%
  bind_rows(all_params %>%
      mutate(config = case_when(!barrier & !boundary ~ "simple",
                                !barrier &  boundary ~ "boundary",
                                barrier & !boundary ~ "barrier")) %>%
      group_by(cutoff, config) %>%
      summarise(n_vertices = first(n_vertices),
                computation_time = first(computation_time),
                .groups = "drop") %>%
      transmute(cutoff, config, n_vertices, 
                metric = "computation_time",
                value = computation_time, 
                conf.low = NA_real_, 
                conf.high = NA_real_)) %>%
  mutate(cutoff = factor(cutoff),
         metric = factor(metric,
                         levels = c("range", "sigma_O", "(Intercept)", "computation_time"),
                         labels = c("spatial range", "sigma O", "intercept", "computation time"))) %>%
  ggplot(aes(x = cutoff, y = value, color = config, group = config)) +
  geom_point(data = ~ dplyr::filter(.x, metric != "computation time"),
             position = position_dodge(0.5), size = 2.5) +
  geom_point(data = ~ dplyr::filter(.x, metric == "computation time"),
             aes(size = n_vertices),
             position = position_dodge(0.5)) +
  geom_errorbar(data = ~ dplyr::filter(.x, metric != "computation time"),
                aes(ymin = conf.low, ymax = conf.high),
                width = 0,position = position_dodge(0.5), na.rm = TRUE) +
  facet_wrap(~metric, scales = "free_y", ncol = 2) +
  scale_size_continuous(name = "Number of vertices", range = c(2, 8)) +
  theme_bw() +
  labs(x = "Cutoff", y = NULL, color = "Meshes configuration")

```
<br/>

```{r echo=FALSE, fig.height=8, fig.width=12, message=FALSE, warning=FALSE}

all_moran %>%
  dplyr::mutate(config = dplyr::case_when(!barrier & !boundary ~ "simple",
                                          !barrier &  boundary ~ "boundary",
                                          barrier & !boundary ~ "barrier"),
                cutoff = factor(cutoff)) %>%
  ggplot2::ggplot(ggplot2::aes(x = year, y = config, fill = (prop_moran_signif*100))) +
  ggplot2::geom_tile(color = "white") +
  ggplot2::geom_text(ggplot2::aes(label = round((prop_moran_signif*100), 3)),
                     color = "black", fontface = "bold", size = 3) +
  ggplot2:: scale_fill_distiller(palette = "RdYlBu")+
  ggplot2::facet_wrap(~cutoff, ncol = 3,
                      labeller = labeller(cutoff = function(x) paste("Cutoff =", x))) +
  ggplot2::theme_bw() +
  ggplot2::labs(x = "years", fill = "% significant \nMoran's index", y = NULL)

```
It shows how the significance of spatial autocorrelation in residuals with Moran’s Index varies with cutoff values and mesh configurations across years. Significant values (* ** ***) indicate remaining spatial structure in residuals (the model does not fully capture spatial patterns), while non-significant results (ns) suggest that spatial autocorrelation is adequately accounted for. The Moran’s Index values indicate the strength and direction of spatial autocorrelation, with values close to zero indicating weak or no spatial structure, positive values indicating clustering of similar residuals, and negative values indicating spatial dispersion.
<br/>  
<br/>

# 3. Plot meshes
<br/>
Simple meshes
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
if (!is.null(plots_meshes_simple)) plots_meshes_simple
```
<br/>  
Boundary meshes
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
if (!is.null(plots_meshes_boundary)) plots_meshes_boundary

```
<br/>  
Barrier meshes
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
if (!is.null(plots_meshes_barrier)) plots_meshes_barrier

```

blue points = centers of triangles located at sea  
red points = centers of triangles positioned on land  
In land triangles, the spatial range is reduced to 10% of the range used at sea (range_fraction = 0.1).
<br/> 
<br/>  


# 3. Prediction maps

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Convert resolution from km to degrees
res_deg_lat <- res_km / 111.32
res_deg_lon <- res_km / (111.32 * cos(50*pi/180))

# Define plot limits with a one-cell buffer
xmin_plot <- range(prediction_grid$lon)[1] - res_deg_lon
xmax_plot <- range(prediction_grid$lon)[2] + res_deg_lon
ymin_plot <- range(prediction_grid$lat)[1] - res_deg_lat
ymax_plot <- range(prediction_grid$lat)[2] + res_deg_lat

land <- ne_download(scale = "large", type = "land", category = "physical", returnclass = "sf") %>%
  st_transform(4326) %>%
  st_crop(xmin = xmin_plot, ymin = ymin_plot, xmax = xmax_plot, ymax = ymax_plot)
  
# Common fill limits across all models
all_est <- c(if(!is.null(predictions_simple)) predictions_simple$est,
             if(!is.null(predictions_barrier)) predictions_barrier$est,
             if(!is.null(predictions_boundary)) predictions_boundary$est)

fill_limits <- range(all_est[is.finite(all_est) & all_est > 0], na.rm = TRUE)

```
<br/>  
Predictions on simple meshes
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}

if(!is.null(predictions_simple)) {
  predictions_simple %>%
    ggplot(aes(lon, lat, fill = est)) +
    geom_raster() +
    geom_sf(data = land, fill = "grey80", inherit.aes = FALSE, color = "grey80") +
    scale_y_continuous(breaks = pretty_breaks(n = 3)) +
    scale_x_continuous(breaks = pretty_breaks(n = 4)) +
    coord_sf(xlim = c(xmin_plot, xmax_plot), ylim = c(ymin_plot, ymax_plot), expand = FALSE) +
    facet_grid(cutoff ~ year) +   
    theme(panel.grid = element_blank(),
          panel.background = element_rect(fill = "white", colour = NA),
          panel.border = element_rect(color = "white", fill = NA, linewidth = 2),
          strip.background = element_rect(fill = "grey40", color = "white"),
          strip.text = element_text(color = "white", face = "bold")) +
    labs(x = "", y = "",  fill = "est") +
    scale_fill_distiller(trans = "log10", 
                         palette = "RdYlBu",
                         limits = fill_limits,
                         labels = label_number(drop0trailing = TRUE))

}
```
<br/>  
Predictions on barrier meshes
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}

if(!is.null(predictions_barrier)) {
  predictions_barrier %>%
    ggplot(aes(lon, lat, fill = est)) +
    geom_raster() +
    geom_sf(data = land, fill = "grey80", inherit.aes = FALSE, color = "grey80") +
    scale_y_continuous(breaks = pretty_breaks(n = 3)) +
    scale_x_continuous(breaks = pretty_breaks(n = 4)) +
    coord_sf(xlim = c(xmin_plot, xmax_plot), ylim = c(ymin_plot, ymax_plot), expand = FALSE) +
    facet_grid(cutoff ~ year) +   
    theme(panel.grid = element_blank(),
          panel.background = element_rect(fill = "white", colour = NA),
          panel.border = element_rect(color = "white", fill = NA, linewidth = 2),
          strip.background = element_rect(fill = "grey40", color = "white"),
          strip.text = element_text(color = "white", face = "bold")) +
    labs(x = "", y = "",  fill = "est") +
    scale_fill_distiller(trans = "log10", 
                         palette = "RdYlBu",
                         limits = fill_limits,
                         labels = label_number(drop0trailing = TRUE))
}
```
<br/>  
Predictions on boundary meshes
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}

if(!is.null(predictions_boundary)) {
  predictions_boundary %>%
    ggplot(aes(lon, lat, fill = est)) +
    geom_raster() +
    geom_sf(data = land, fill = "grey80", inherit.aes = FALSE, color = "grey80") +
    scale_y_continuous(breaks = pretty_breaks(n = 3)) +
    scale_x_continuous(breaks = pretty_breaks(n = 4)) +
    coord_sf(xlim = c(xmin_plot, xmax_plot), ylim = c(ymin_plot, ymax_plot), expand = FALSE) +
    facet_grid(cutoff ~ year) +   
    theme(panel.grid = element_blank(),
          panel.background = element_rect(fill = "white", colour = NA),
          panel.border = element_rect(color = "white", fill = NA, linewidth = 2),
          strip.background = element_rect(fill = "grey40", color = "white"),
          strip.text = element_text(color = "white", face = "bold")) +
    labs(x = "", y = "",  fill = "est") +
    scale_fill_distiller(trans = "log10", 
                         palette = "RdYlBu",
                         limits = fill_limits,
                         labels = label_number(drop0trailing = TRUE))
}
```




