---
title: "Mesh sensitivy analysis per species and region"
date: "`r Sys.Date()`"
output: 
  rmdformats::downcute:
  highlight: zenburn
html_document :
  toc: true
toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = getwd())                                # garde le working dir actif
knitr::opts_knit$set(envir = globalenv())                               # utilise l'env global
knitr::opts_chunk$set(echo = FALSE, message=FALSE,warning=FALSE)
```


<br/>
 
```{r echo=FALSE}

print(sp_scientific)

```
<br/>
<br/>

# 1. Sanity results for meshes configurations
<br/>
```{r echo=FALSE, message=FALSE, warning=FALSE}

conv_simple <- if (!is.null(sanity_simple)) names(sanity_simple$cutoffs_converged_models) else character(0)
fail_simple <- if (!is.null(sanity_simple)) names(sanity_simple$cutoffs_sanity_failed) else character(0)

conv_barrier <- if (!is.null(sanity_barrier)) names(sanity_barrier$cutoffs_converged_models) else character(0)
fail_barrier <- if (!is.null(sanity_barrier)) names(sanity_barrier$cutoffs_sanity_failed) else character(0)

conv_boundary <- if (!is.null(sanity_boundary)) names(sanity_boundary$cutoffs_converged_models) else character(0)
fail_boundary <- if (!is.null(sanity_boundary)) names(sanity_boundary$cutoffs_sanity_failed) else character(0)

conv_simple <- if (!is.null(sanity_simple_no_max)) names(sanity_simple_no_max$cutoffs_converged_models) else character(0)
fail_simple <- if (!is.null(sanity_simple_no_max)) names(sanity_simple_no_max$cutoffs_sanity_failed) else character(0)

conv_barrier <- if (!is.null(sanity_barrier_no_max)) names(sanity_barrier_no_max$cutoffs_converged_models) else character(0)
fail_barrier <- if (!is.null(sanity_barrier_no_max)) names(sanity_barrier_no_max$cutoffs_sanity_failed) else character(0)

conv_boundary <- if (!is.null(sanity_boundary_no_max)) names(sanity_boundary_no_max$cutoffs_converged_models) else character(0)
fail_boundary <- if (!is.null(sanity_boundary_no_max)) names(sanity_boundary_no_max$cutoffs_sanity_failed) else character(0)


sanity_results_df <- dplyr::bind_rows(

  tibble(cutoff = west_params$with_max$cutoff_values, 
         config = "simple outer mesh", 
         status = dplyr::case_when(
           paste0("cutoff_", cutoff) %in% conv_simple ~ "converged",
           paste0("cutoff_", cutoff) %in% fail_simple ~ "failed")),
  tibble(cutoff = west_params$with_max$cutoff_values, 
         config = "barrier outer mesh",
         status = dplyr::case_when(
           paste0("cutoff_", cutoff) %in% conv_barrier ~ "converged",
           paste0("cutoff_", cutoff) %in% fail_barrier ~ "failed")),
  tibble(cutoff = west_params$with_max$cutoff_values,
         config = "boundary outer mesh",
         status = dplyr::case_when(
           paste0("cutoff_", cutoff) %in% conv_boundary ~ "converged",
           paste0("cutoff_", cutoff) %in% fail_boundary ~ "failed")),
    tibble(cutoff = west_params$no_max$cutoff_values, 
         config = "simple", 
         status = dplyr::case_when(
           paste0("cutoff_", cutoff) %in% conv_simple ~ "converged",
           paste0("cutoff_", cutoff) %in% fail_simple ~ "failed")),
  tibble(cutoff = west_params$no_max$cutoff_values, 
         config = "barrier",
         status = dplyr::case_when(
           paste0("cutoff_", cutoff) %in% conv_barrier ~ "converged",
           paste0("cutoff_", cutoff) %in% fail_barrier ~ "failed")),
  tibble(cutoff = west_params$no_max$cutoff_values,
         config = "boundary",
         status = dplyr::case_when(
           paste0("cutoff_", cutoff) %in% conv_boundary ~ "converged",
           paste0("cutoff_", cutoff) %in% fail_boundary ~ "failed"))
  
  )

sanity_results_df$config <- factor(sanity_results_df$config,
                                   levels = c("boundary","barrier","simple", 
                                              "boundary outer mesh",
                                              "barrier outer mesh", 
                                              "simple outer mesh"))

ggplot(sanity_results_df, aes(x = factor(cutoff), y = config, fill = status)) +
  geom_tile(color = "white") +
  scale_fill_manual(values = c(converged = "#2ca25f", failed = "#ffb703")) +
  labs(x = "Cutoff", y = "Configuration", fill = "Model status") +
  theme_minimal()

```

<br/>  
<br/>

# 2. Estimates and metrics for meshes configurations
<br/>

```{r echo=FALSE, fig.height=8, fig.width=12, message=FALSE, warning=FALSE}

plot_df <- bind_rows(
  all_params %>%
    transmute(cutoff, config, metric = term, value = estimate, conf.low, conf.high),
  all_params %>%
    distinct(cutoff, config, cAIC, n_vertices) %>%
    pivot_longer(cols = c(cAIC, n_vertices), names_to = "metric", values_to = "value") %>%
    mutate(conf.low = NA, conf.high = NA)) %>%
  mutate(cutoff = factor(cutoff), metric = factor(metric,
      levels = c("(Intercept)", "range", "sigma_O", "cAIC", "n_vertices"),
      labels = c("intercept", "spatial range", "sigma O", "cAIC", "number of vertices")))

ggplot(plot_df, aes(x = cutoff, y = value, color = config, group = config)) +
  geom_point(position = position_dodge(0.5), size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                width = 0, position = position_dodge(0.5), na.rm = TRUE) +
  facet_wrap(~metric, scales = "free_y") +
  theme_bw() +
  labs(x = "Cutoff", y = NULL, color = "Configuration")

```

<br/>  
<br/>

# 3. Plot meshes
<br/>
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
if (!is.null(plots_meshes_simple)) plots_meshes_simple
```
<br/>  
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
if (!is.null(plots_meshes_boundary)) plots_meshes_boundary
```
<br/>  
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
if (!is.null(plots_meshes_barrier)) plots_meshes_barrier
```
<br/>  
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
if (!is.null(plots_meshes_simple_no_max)) plots_meshes_simple_no_max
```
<br/>  
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
if (!is.null(plots_meshes_boundary_no_max)) plots_meshes_boundary_no_max
```
<br/>  
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
if (!is.null(plots_meshes_barrier_no_max)) plots_meshes_barrier_no_max
```
blue points = centers of triangles located at sea  
red points = centers of triangles positioned on land  
In land triangles, the spatial range is reduced to 10% of the range used at sea (range_fraction = 0.1).
<br/> 
<br/>  


# 3. Coefficient of variarion

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Convert resolution from km to degrees
  res_deg_lat <- res_km / 111.32
  res_deg_lon <- res_km / (111.32 * cos(50*pi/180))

  # Define plot limits with a one-cell buffer
  xmin_plot <- range(prediction_grid$lon)[1] - res_deg_lon
  xmax_plot <- range(prediction_grid$lon)[2] + res_deg_lon
  ymin_plot <- range(prediction_grid$lat)[1] - res_deg_lat
  ymax_plot <- range(prediction_grid$lat)[2] + res_deg_lat

  land <- ne_download(scale = "large", type = "land", category = "physical", returnclass = "sf") %>%
    st_transform(4326) %>%
    st_crop(xmin = xmin_plot, ymin = ymin_plot, xmax = xmax_plot, ymax = ymax_plot)
```

Simple mesh 
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}

if(!is.null(cv_outputs_simple_no_max)) {
  cv_outputs_simple_no_max %>%
    filter(is.finite(cv), cv > 0) %>%
    ggplot(aes(lon, lat, fill = cv)) +
    geom_raster() +
    geom_sf(data = land, fill = "grey80", inherit.aes = FALSE, color = "grey80") +
    scale_y_continuous(breaks = pretty_breaks(n = 3)) +
    scale_x_continuous(breaks = pretty_breaks(n = 4)) +
    coord_sf(xlim = c(xmin_plot, xmax_plot), ylim = c(ymin_plot, ymax_plot), expand = FALSE) +
    facet_grid(cutoff ~ year) +   
    theme(panel.grid = element_blank(),
          panel.background = element_rect(fill = "white", colour = NA),
          panel.border = element_rect(color = "white", fill = NA, linewidth = 2),
          strip.background = element_rect(fill = "grey40", color = "white"),
          strip.text = element_text(color = "white", face = "bold")) +
    labs(x = "", y = "",  fill = "CV") +
    scale_fill_viridis_c(trans  = "log10",
                         labels = label_number(drop0trailing = TRUE))
}
```

Barrier mesh with outer mesh
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}

if(!is.null(cv_outputs_barrier_no_max)) {
  cv_outputs_barrier_no_max %>%
    filter(is.finite(cv), cv > 0) %>%
    ggplot(aes(lon, lat, fill = cv)) +
    geom_raster() +
    geom_sf(data = land, fill = "grey80", inherit.aes = FALSE, color = "grey80") +
    scale_y_continuous(breaks = pretty_breaks(n = 3)) +
    scale_x_continuous(breaks = pretty_breaks(n = 4)) +
    coord_sf(xlim = c(xmin_plot, xmax_plot), ylim = c(ymin_plot, ymax_plot), expand = FALSE) +
    facet_grid(cutoff ~ year) +   
    theme(panel.grid = element_blank(),
          panel.background = element_rect(fill = "white", colour = NA),
          panel.border = element_rect(color = "white", fill = NA, linewidth = 2),
          strip.background = element_rect(fill = "grey40", color = "white"),
          strip.text = element_text(color = "white", face = "bold")) +
    labs(x = "", y = "",  fill = "CV") +
    scale_fill_viridis_c(trans  = "log10",
                         labels = label_number(drop0trailing = TRUE))
}
```

Boundary mesh with outer mesh
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}

if(!is.null(cv_outputs_boundary_no_max)) {
  cv_outputs_boundary_no_max %>%
    filter(is.finite(cv), cv > 0) %>%
    ggplot(aes(lon, lat, fill = cv)) +
    geom_raster() +
    geom_sf(data = land, fill = "grey80", inherit.aes = FALSE, color = "grey80") +
    scale_y_continuous(breaks = pretty_breaks(n = 3)) +
    scale_x_continuous(breaks = pretty_breaks(n = 4)) +
    coord_sf(xlim = c(xmin_plot, xmax_plot), ylim = c(ymin_plot, ymax_plot), expand = FALSE) +
    facet_grid(cutoff ~ year) +   
    theme(panel.grid = element_blank(),
          panel.background = element_rect(fill = "white", colour = NA),
          panel.border = element_rect(color = "white", fill = NA, linewidth = 2),
          strip.background = element_rect(fill = "grey40", color = "white"),
          strip.text = element_text(color = "white", face = "bold")) +
    labs(x = "", y = "",  fill = "CV") +
    scale_fill_viridis_c(trans  = "log10",
                         labels = label_number(drop0trailing = TRUE))
}
```

Simple mesh with outer mesh
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}

if(!is.null(cv_outputs_simple)) {
  cv_outputs_simple %>%
    filter(is.finite(cv), cv > 0) %>%
    ggplot(aes(lon, lat, fill = cv)) +
    geom_raster() +
    geom_sf(data = land, fill = "grey80", inherit.aes = FALSE, color = "grey80") +
    scale_y_continuous(breaks = pretty_breaks(n = 3)) +
    scale_x_continuous(breaks = pretty_breaks(n = 4)) +
    coord_sf(xlim = c(xmin_plot, xmax_plot), ylim = c(ymin_plot, ymax_plot), expand = FALSE) +
    facet_grid(cutoff ~ year) +   
    theme(panel.grid = element_blank(),
          panel.background = element_rect(fill = "white", colour = NA),
          panel.border = element_rect(color = "white", fill = NA, linewidth = 2),
          strip.background = element_rect(fill = "grey40", color = "white"),
          strip.text = element_text(color = "white", face = "bold")) +
    labs(x = "", y = "",  fill = "CV") +
    scale_fill_viridis_c(trans  = "log10",
                         labels = label_number(drop0trailing = TRUE))
}
```

Barrier mesh with outer mesh
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}

if(!is.null(cv_outputs_barrier)) {
  cv_outputs_barrier %>%
    filter(is.finite(cv), cv > 0) %>%
    ggplot(aes(lon, lat, fill = cv)) +
    geom_raster() +
    geom_sf(data = land, fill = "grey80", inherit.aes = FALSE, color = "grey80") +
    scale_y_continuous(breaks = pretty_breaks(n = 3)) +
    scale_x_continuous(breaks = pretty_breaks(n = 4)) +
    coord_sf(xlim = c(xmin_plot, xmax_plot), ylim = c(ymin_plot, ymax_plot), expand = FALSE) +
    facet_grid(cutoff ~ year) +   
    theme(panel.grid = element_blank(),
          panel.background = element_rect(fill = "white", colour = NA),
          panel.border = element_rect(color = "white", fill = NA, linewidth = 2),
          strip.background = element_rect(fill = "grey40", color = "white"),
          strip.text = element_text(color = "white", face = "bold")) +
    labs(x = "", y = "",  fill = "CV") +
    scale_fill_viridis_c(trans  = "log10",
                         labels = label_number(drop0trailing = TRUE))
}
```

Boundary mesh with outer mesh
```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}

if(!is.null(cv_outputs_boundary)) {
  cv_outputs_boundary %>%
    filter(is.finite(cv), cv > 0) %>%
    ggplot(aes(lon, lat, fill = cv)) +
    geom_raster() +
    geom_sf(data = land, fill = "grey80", inherit.aes = FALSE, color = "grey80") +
    scale_y_continuous(breaks = pretty_breaks(n = 3)) +
    scale_x_continuous(breaks = pretty_breaks(n = 4)) +
    coord_sf(xlim = c(xmin_plot, xmax_plot), ylim = c(ymin_plot, ymax_plot), expand = FALSE) +
    facet_grid(cutoff ~ year) +   
    theme(panel.grid = element_blank(),
          panel.background = element_rect(fill = "white", colour = NA),
          panel.border = element_rect(color = "white", fill = NA, linewidth = 2),
          strip.background = element_rect(fill = "grey40", color = "white"),
          strip.text = element_text(color = "white", face = "bold")) +
    labs(x = "", y = "",  fill = "CV") +
    scale_fill_viridis_c(trans  = "log10",
                         labels = label_number(drop0trailing = TRUE))
}
```

```{r echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}

all_cv_outputs %>%
    mutate(config = factor(config, levels = c("simple", "barrier", "boundary", "simple outer mesh",
                                              "barrier outer mesh", "boundary outer mesh"))) %>%
  ggplot2::ggplot(aes(x = factor(cutoff), y = cv, color = factor(cutoff)))+
  ggplot2::geom_boxplot()+
  ggplot2::geom_violin(fill=NA)+
  ggplot2::geom_hline(yintercept = 1, linetype = "dashed", color = "grey70")+
  ggplot2::geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey70")+
  ggplot2::scale_color_brewer(palette = "Dark2")+
  ggplot2::facet_wrap(~config)+
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "none")
  

```

